---
output: github_document
---

<!-- README.md is generated from README.rmd. Please edit that file -->

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
# devtools::load_all()
library(plt)
```

# plt

Persistence landscapes are a vectorization of persistence data (also called persistence diagrams), [originated by Peter Bubenik](https://jmlr.csail.mit.edu/papers/v16/bubenik15a.html), that have useful statistical properties including linearity and an inner product.
This is an R package interface to [Paweł Dłotko's Persistence Landscapes Toolbox](https://www2.math.upenn.edu/~dlotko/persistenceLandscape.html), [developed with Bubenik](https://www.sciencedirect.com/science/article/pii/S0747717116300104) to efficienctly compute and calculate with persistence landscapes.
It was adapted from [Jose Bouza's **tda-tools** package](https://github.com/jjbouza/tda-tools), a TDA pipeline used by [Bubenik's lab](https://people.clas.ufl.edu/peterbubenik/researchgroup/).

## Installation

Since the package is not on CRAN, we suggest that you install **remotes** on your system. To do this just open an R session and run:

```{r, eval=FALSE}
install.packages("remotes")
```

Then install the package from the GitHub repository as follows:

```{r, eval=FALSE}
remotes::install_github("corybrunson/plt")
```

Alternatively, you can clone or download the code repository and install the package from source (from within the directory):

```{r, eval=FALSE}
devtools::install()
```

You should now be able to load the package normally from an R session using
`library("plt")`.

## Quickstart Guide

The **plt** package supports various operations involving persistence landscapes:

* Compute persistence landscapes from persistence data
* Perform Hilbert space operations (vector space operations and an inner product) on persistence landscapes
* Plot persistence landscapes

Examples and tests in **plt** rely on other packages to simulate data and to compute persistence diagrams from data:

* **tdaunif** provides functions to sample uniformly from various immersed manifolds.
* **ripserr** and **TDA** provide functions to compute persistence data from point clouds and distance matrices.

### Calculation

**plt** introduces the 'Rcpp_PersistenceLandscape' S4 class, which is exposed using **Rcpp** from the underlying 'PersistenceLandscape' C++ class.
Instances of this class can be created using `new()` but the recommended way is to use `landscape()`.
This function accepts either a single matrix of persistence data or a persistence diagram object, which is a specially formatted list containing the first and most important element `$pairs`.
The `$pairs` entry is itself a list, of a 2-column matrix of persistence pairs for each homological degree from 0 (`$pairs[[1]]`) to the maximum degree calculated.
The generic conversion function `as_persistence()` includes methods for outputs from `ripserr::vietoris_rips()` and from `TDA::*Diag()`.

To begin an illustration, we noisily sample 60 points from a figure eight and compute the persistence diagram of the point cloud:

```{r}
pc <- tdaunif::sample_lemniscate_gerono(60, sd = .1)
pd <- ripserr::vietoris_rips(pc, max_dim = 1, threshold = 2, p = 2)
print(pd)
```

We the convert the persistence data to the preferred persistence diagram format and inspect some of its features:

```{r}
pd <- as_persistence(pd)
print(head(pd$pairs[[1]]))
print(head(pd$pairs[[2]]))
```

This allows us to compute a persistence landscape---in this case, for the 1-dimensional features.
Here we compute the landscape exactly, which can be cost-prohibitive for larger persistence data:

```{r}
pl1 <- landscape(pd, degree = 1, exact = TRUE)
print(pl1)
```

### Class

The object `pl1` is not an array, but rather an object that encapsulates both the data that encode a landscape and several basic operations that can be performed on it.
This allows us to work with persistence landscapes without worrying about pre-processing their representations.
At any point, the underlying landscape can be extracted using `$getInternal()`, which in the case of an exactly calculated landscape returns a list of 2-column matrices, each matrix containing coordinates that define one level of the landscape as a piecewise linear function:

```{r}
print(length(pl1$getInternal()))
print(pl1$getInternal())
```

The length of this list is the number of levels of the landscape.

An alternative, approximate construction computes the value of each level of the landscape at each point on a 1-dimensional grid, ranging from `min_b` to `max_b` at increments of `by`.
A landscape constructed using a discrete approximation is stored as a 3-dimensional array of dimensions (levels, values, 2), with one level per feature (some of which may be trivial) and one value per grid point, stored as $x,y$ pairs along the third dimension.

```{r}
pl1d <- landscape(pd, degree = 1, min_b = 0, max_b = 2, by = 0.01)
print(dim(pl1d$getInternal()))
print(pl1d$getInternal())
```

Exactly computed landscapes can be converted to discrete landscape objects, but the other direction is not well-defined:

```{r}
pl1$getDiscrete()
try(pl1d$getExact())
```

### Visualization

**plt** provides a `plot()` method for the 'Rcpp_PersistenceLandscape' class.
It uses **grDevices** to build color palettes, and as such its default palette is viridis; but the user may supply the name of a recognized palette or a sequence of colors between which to interpolate:

```{r}
plot(pl1, palette = "terrain")
plot(pl1d, palette = "terrain")
```

### Vector Operations



### Inner Product


